---
title: 'Lecture 3: Data Visualization'
author: "<br> <br >Yanfei Kang <br> yanfeikang@buaa.edu.cn"
date: "School of Economics and Management <br> Beihang University"
output:
  slidy_presentation:
  footer: "Lecture 3: Data Visualization"
css: ../styles/ykstyle.css
logo: buaalogo.png
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# What we are going to learn today ...

1. basic graphics with **R**
2. elegant graphics with **ggplot2**
3. interactive plotting with **manipulate**
3. interactive graphs with **plotly**
4. interactive web apps with **shiny**


# {.alert}
<br>
<br>
Basic graphics with **manipulate**

# Basic graphics

```{r}
plot(cars)
```

#### Check this out:
```{r, eval=FALSE}
library(manipulate)
manipulate(
  plot(cars, xlim=c(0,x.max)),  
  x.max=slider(15,25))
```

# Basic graphics

```{r}
plot(longley[, 'Year'], longley[, 'GNP'], type = 'l', main = 'GNP')
```

```{r, eval=FALSE}
manipulate(
  plot(longley[, 'Year'], longley[,factor], type = 'l', main = factor),
  factor = picker("GNP", "Unemployed", "Employed"))
```

# Combine controls

```{r}
manipulate(
  plot(longley[, 'Year'], longley[,factor], xlim = c(1947, x.max), main = factor, ann = label, type = 'l'),
   x.max = slider(1956, 1962, step=1, initial = 1962),
  factor = picker("GNP", "Unemployed", "Employed"),
  label = checkbox(TRUE, "Draw Labels"))
```

# References

1. Philipp K. Janert (2011). Data analysis with open source tools. O'Reilly.
2. Joseph Adler (2012). R in a Nutshell (2ed). O'Reilly.
3. [R for data science](http://r4ds.had.co.nz/index.html)
