---
title: 'Lecture 8: Rmarkdown - a tool for reporting results from R'
author: "Yanfei Kang"
output:
  html_document:
    highlight: tango
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
bibliography: L8.bib
---


# Packages required in this lecture

```
install.packages(c("rmarkdown", "forecast"))
```


# Why R markdown?

- Much **easier** *syntax* ~~syntax~~ than LaTex or Html
- Dynamic: easy to update and work with the R codes
- Multiple output formats
- Makes presentation easy
- Keep me organized of weekly research progress reports


# Header 1
## Header 2


# Creation of lists
## unordered
- Item 1
- Item 2

## ordered
1. Item 1
2. Item 2
    + item 2a
    + item 2b


# Inline codes
Write inline **R** code using backtick quotes: `forecast()`


# Blockqotes
> It's always better to give than to receive.


# Equations

- $x + y = z$ for inline equations
- $$ x + y = z $$ 


# Hyperlink

[R markdown tutorial](http://rmarkdown.rstudio.com)

************************************


# R code chunks with plot 

See @hyndman2014forecasting.

```{r RcodeDemoPlot}
library(forecast)
plot(stlf(AirPassengers, lambda=0))
```


# R code chunks with table
```{r RcodeDemoTable}
knitr::kable(head(iris))
```



# Plain code block 

```
library(forecast)
plot(stlf(AirPassengers, lambda=0))
```


# References




