---
title: 'Lecture 1: R Tips'
author: "<br> <br >Yanfei Kang <br> yanfeikang@buaa.edu.cn"
date: "School of Economics and Management <br> Beihang University"
output:
  slidy_presentation:
    footer: "Lecture 1: R Tips"
    css: ../styles/ykstyle.css
logo: buaalogo.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```
# In this lecture we are going to learn ...

- How to install and use **R**
- **R** coding style 
- some nice **R** tips


# Packages required in this lecture
```{r}
install.packages(c("forecast", "sos", "formatR"))
```

# {.alert}
<br>
<br>
How to install and use **R**
                  
# Getting and Installing R

- Totally free!
- Download R on its [official website](https://www.r-project.org).


# R User Interface

- Direct **R**
- Rstudio
    - One of the most popular ways to run R.
    - Free, open-source integrated development environment (IDE) for R.
    - Many additional fantastic features.
- Command line in Linux and Unix.



# Editor

- What editor do you usually use?
- Use a good text editor such as vim, sublime text, text wrangler, notepad, etc
- With syntax highlighting, otherwise, it's hard to detect errors
- Or use an Integrated Development Environment (IDE) like RStudio

# Use an IDE: Rstudio

- Syntax highlighting
- Able to evaluate R code
      + by line
      + by selection 
      + entire file
- Command auto-completion

# {.alert}
<br>
<br>
**R** coding style

# File names 

- File names should end in .R and, of course, be *meaningful*. 
- GOOD: `predict_ad_revenue.R` 
- BAD: `foo.R`

# Choose the names of variables and functions carefully

#### What we should do

- The preferred form for variable names is all lower case letters and words separated with dots (variable.name), but variableName is also accepted. Generally, variable names should be nouns.
    - GOOD: avg.clicks 
    - OK: avgClicks 
    - BAD: avg_Clicks
- Function names have initial capital letters and no dots. Function names are mostly verbs.
    - GOOD: CalculateAvgClicks 
    - BAD: calculate_avg_clicks , calculateAvgClicks 
- Choose a consistent naming style

#### What we should not do

- Don't use underscores (_) or hyphens (-).
- Avoid using names of existing functions and variables like `mean', `median' etc.
- Avoid using meaningless names like a, b, c, ..., aa, bb, cc, ...









# White Spaces

- around operators (=, +, -, <-, etc)
- put a space after a comma, and never before

```{r}
x <- c(1:10)
x.average<-mean(x,na.rm=TRUE)
```

$\Rightarrow$

```{r}
x.average <- mean(x, na.rm = TRUE)
```

- split long lines at meaningful places

Don’t be afraid of splitting one long line into individual pieces!

```{r}
n <- matrix(sample(1:100, 9), nrow = 3, ncol = 3, byrow = TRUE)
```

$\Rightarrow$

```{r}
n <- matrix(sample(1:100, 9),
           nrow = 3,
           ncol = 3,
           byrow = TRUE)
```

# Curly braces
- An opening curly brace should never go on its own line and should always be followed by a new line. 
- A closing curly brace should always go on its own line, unless it’s followed by else.
- Always begin the body of a block on a new line.
- Always indent the code inside curly braces.

```{r}
if (y < 0) {print("y is negative")}
```
$\Rightarrow$
```{r}
if (y < 0) {
  print("y is negative")
}
```
# Indenting 

- Use two spaces
- Can help in detecting errors in your code because it can expose lack of symmetry
- Reindenting using RStudio


```{r}
if (y < 0) {
print("y is negative")
}
```
$\Rightarrow$
```{r}
if (y < 0) {
  print("y is negative")
}
```

# Make your code tidy in a second!

- Reformat and reindent in Rstudio.
- **formatR** package in **R**. You can even make a folder of `.R` files tidy using `tidy.dir()`.



# Header, Line spaces and Comments

- Add a Header for your file
- Add lots of comments
- Use blank lines to separate blocks of code and comments to say what the block does. Remember that in a few months, you may not follow your own code any better than a stranger.


```{r}
x <- c(1:10)
x.mean = mean(x)
x.var = var(x)
```

$\Rightarrow$

```{r}
# =============================================
# Title
# Author: Yanfei Kang
# Date: Mar 23, 2017
# Description: your purpose
# =============================================

x <- c(1:10)

# getting the mean of x
x.mean = mean(x)

# getting the variance of x
x.var = var(x)
```

# Function Documentation

- Functions should contain a comments section immediately below the function definition line. 
- These comments should include
    - a one-sentence description of the function
    - a list of the function's arguments, denoted by Args:, with a description of each (including the data type)
    - a description of the return value, denoted by Returns:. 
    - The comments should be descriptive enough that a caller can use the function without reading any of the function's code.

```{r}
CalculateSampleCovariance <- function(x, y, verbose = TRUE) {
  # Computes the sample covariance between two vectors.
  #
  # Args:
  #   x: One of two vectors whose sample covariance is to be calculated.
  #   y: The other vector. x and y must have the same length, greater than one,
  #      with no missing values.
  #   verbose: If TRUE, prints sample covariance; if not, not. Default is TRUE.
  #
  # Returns:
  #   The sample covariance between x and y.
  n <- length(x)
  # Error handling
  if (n <= 1 || n != length(y)) {
    stop("Arguments x and y have different lengths: ",
         length(x), " and ", length(y), ".")
  }
  if (TRUE %in% is.na(x) || TRUE %in% is.na(y)) {
    stop(" Arguments x and y must not have missing values.")
  }
  covariance <- var(x, y)
  if (verbose)
    cat("Covariance = ", round(covariance, 4), ".\n", sep = "")
  return(covariance)
}
```

# {.alert}
<br>
<br>
Some nice  **R** tips

# How to find the right function 

- Functions in installed packages

```{r}
library(forecast)
help.search("auto.arima")
??auto.arima
```
- Functions in other CRAN packages
 
```{r}
library(sos)
findFn("arima")
RSiteSearch("arima")
```
 
# Digging into functions

- Type `?sort` for the usage of the function `sort()`.
- Typing the name of a function gives its definition.
- Type `forecast:::estmodel` for hidden functions.
- Download the tar.gz file from CRAN if you want to see any underlying **C** or **Fortran** code.

# Organize your R projects

#### Basic idea

- Every paper, book or scientific report is a 'project'.
- Every project has its own folder and R workspace.
- Every project is entirely scripted. That is, all analysis, graphs and tables must be able to be generated by running one script. 
    + This script sources all other R files in the correct order and yields all the required results. This script could be in `main.R` or `main.Rmd`.
    + `functions.R` contains all non-packaged functions used in the project.
    + each function can not be too long.
    
  

# Look at other people's codes 
- https://github.com/hadley
- https://github.com/yihui
- https://github.com/karthik
- https://github.com/kbroman
- https://github.com/cboettig
- https://github.com/garrettgman

# Getting help
- For programming questions: StackOverflow.com
- For statistical questions: CrossValidated.com

# Keep up-to-date
- RStudio blog: blog.rstudio.org
- R-bloggers: www.r-bloggers.com
- It takes time to develop your own style. Once it is developed, it is really hard to be changed. So please be careful at the beginning. 


    
# Your turn (1) 💓

- Write a function to solve the roots of given quadratic equation $ax^2 + bx + c = 0$ with $a$, $b$ and $c$ as input arguments. 

- Test your function on some simple equations.

- Keep in mind the styles we have learnt.


# Your turn (2) 💓

- Write two messy `.R` files and put them in a folder.
- Use `tidy.dir()` to make them tidy.

# References

- [Google R style guide](https://google.github.io/styleguide/Rguide.xml)
- [Rob's tips](http://robjhyndman.com/hyndsight/simpler/)

# Further readings

- [Hadley's style](http://adv-r.had.co.nz/Style.html)